---
title: "13th"
author: "enrique95"
date: '`r format(Sys.time(), "%m/%Y")`'
output: md_document
---

# Dedication

I would like to thank Iván Logrosán for the initial idea, the UC3M for helping me through this journey of learning, the NHGIS workers for their help and resources and last but not least, the discoRd server, the aces of R in discord.


# 1. Introduction

## 1.1 Motivation of work and goals

Slavery has been a very controversial topic that even nowadays has ramifications in many different sectors and markets, be "legal" but out of public scrutiny, or illegal, yet scarcely prosecuted. This project aims to analyze the effect of slavery and its abolition with the 13th Amendment on the agriculture through the use of agricultural, demographic and geographical data. The goal of this project is to analyze just the economic background. From the ethical background there is not much discussion regarding this topic than what it is already established for years.

## 1.2 Resources and data sources

The data has been retrieved from the NHGIS database and analyzed through the `ipumsr` package, and wrangled with `base` R and `tidyverse` functions. The prices of many agricultural products that are not inside the NHGIS database come directly from the US Census.

## 1.3 Metodology

The data will be analyzed through a Diff-in-diff model. This model is ideal for analyzing changes of trend when a policy is enforced or some major event in a single point in time is potentially responsible of the change of pace of some activity. In this case the 13th Amendment is this policy that forbade the use of slave work in the USA. Another major event, intrinsically related to the 13th Amendment and of great impact in the US was the Civil War. This will be analyzed too.

# 2. Getting started
## 2.1 Packages

First the packages necessary for reading and managing the data are installed and loaded. For dependency issues in Linux check the `linuxdep.sh` script

```{r}
# Package installation
install.packages("plm") #for panel data use
install.packages("stargazer") #for building tables
install.packages("robustbase") #for heteroskedastic errors
install.packages("tidyverse") #for data wrangling
install.packages("lmtest") #for testing heteroskedasticity 
install.packages("ipumsr") #for loading NHGIS data
install.packages("rmarkdown") #for writing this .rmd file
install.packages("tesseract") #for reading images into text
install.packages("pdftools") #for uploading pdf as image
install.packages("sf") #for spatial data

library(sf)
library(plm)
library(stargazer)
library(pdftools)
library(rmarkdown)
library(robustbase)
library(tidyverse)
library(lmtest)
library(ipumsr)
library(tesseract)

```

## 2.2 Loading the data

First the relative paths are added so the `ipumsr` functions know the zip files from which can extract the data

```{r}
pathshp <- "nhgis0027_shape.zip"
pathcsv <- "nhgis0027_csv.zip"
```
